---
title: "Analysis of SARS-CoV2 cases in a pediatric hospital"
author: "Marta Mainetti"
date: "`r date()`"
format: 
  html:
    toc: true
    code-fold: show
    code-tools: true
    df-print: paged
---


# Import libraries and prepare setting

```{r, warning = FALSE, message = FALSE}
library(broom)
library(cowplot)
library(dplyr)
library(ggplot2)
library(ggpointdensity)
library(gtsummary)
library(medicaldata)
library(skimr)
library(unibeCols)
```

The aim of this work is to evaluate the relationship between the incidence of 
cases of covid 19 (measured with the positivity of the covid test) and the time since the start of the pandemic. 
For this purpose, I will use the data `covid_testing` from the library
`medicaldata`, that is a dataset containing details of SARS-CoV-2 testing in 2020 at the Children's Hospital of Philadelphia, Pennsylvania (CHOP).

# Load data

```{r}
covid_testing_raw <- medicaldata::covid_testing
```

We proceed with an initial data manipulation. We exclude the subjects with an invalid
result in the test, and create a binary outcome: 

```{r}
covid_testing <- covid_testing_raw %>% 
  filter(result != "invalid") %>% 
  mutate(outcome = as.numeric(result == "positive")) %>% 
  distinct()

head(covid_testing)
```

As we can see, we have the column `subject_id` that identifies patients in the
dataset. Let's check if we have multiple measurements per subject:

```{r}
length(unique(covid_testing$subject_id)) ; nrow(covid_testing)
```

Since the number of records is larger than the unique values of `subject_id`, this means that some subjects have more than one measurement. How should we deal with that? We have some options:

- We ignore it. It is not relevant to the analysis whether we have two different testings

- We keep either the first or the last, considering that there might be correlations between the events (e.g. subjects that have already been positive have higher probability of compared to children that have never been tested positive)

I decided to go with the second option and keep only the results of the first covid test performed on each child:

```{r}
covid_testing <- covid_testing %>% 
  arrange(subject_id, pan_day) %>% 
  group_by(subject_id) %>% 
  slice(1) %>% 
  ungroup()
```


# Preliminary analysis

## Descriptive statistics

In this section I will perform some descriptive analysis of the dataset. 

```{r}
skim_without_charts(covid_testing[, sapply(covid_testing, is.numeric)])
```

And we should also see how many positive cases we have out of the total:

```{r}
t <- table(covid_testing$outcome); t
round(t/ nrow(covid_testing), 3)
```


There is a huge imbalance between the two groups, and this should be considered
when implementing the analysis. 


## Data visualization

It is useful to visualize the number of cases overtime, together with the correlation

```{r}
g <- covid_testing %>% group_by(pan_day) %>% 
  summarize(cases_per_day = sum(result == "positive"))

cor.time.cases <- round(cor(g$pan_day, g$cases_per_day, method = "spearman"), 3)

ggplot(data = g, aes(x = pan_day, y = cases_per_day)) +
  geom_line() +
  annotate("text", x = 7, y = 15, label = paste("r =", cor.time.cases)) +
  labs(x = "Days from start of the pandemic",
       y = "Daily incidence")
```

Now, let's check the differences in the two groups, i.e. distinguishing the subjects who were
tested positive and negative:

```{r}
ggplot(data = covid_testing, aes(x = age, y = result, fill = result)) +
  geom_boxplot() +
  scale_fill_manual(breaks = c("positive", "negative"),
                    values = c(unibeRedS()[2], unibeMustardS()[2])) +
  theme(legend.position = "none")
```

We can see that age seems larger in positive subjects, compared to negative, and
that we do not have data only related to children but also to adults (even if they
are outliers).

Now, let' see how days since pandemic are related to age: 

```{r}
p1 <- ggplot(data = covid_testing, aes(x = pan_day, y = age, col = result)) +
  geom_point(size = 1) +
  scale_color_manual(breaks = c("positive", "negative"),
                    values = c(unibeRedS()[1], unibeMustardS()[2])) +
  labs(x = "Days after start of pandemic", y = "Age", 
       col = "Test result") +
  theme(legend.position = "bottom")

p2 <- ggplot(data = covid_testing, aes(x = pan_day, y = age)) +
  geom_pointdensity(size = 1) +
  labs(x = "Days after start of pandemic", y = "Age") +
  theme(legend.position = "bottom")

cowplot::plot_grid(plotlist = list(p1, p2), align = "h", labels = c("A", "B"))
```

As we can see, plot A displays different colours for positive and negative,
while plot B shows the density of points. We can say that most subjects are in
the age range (0, 25)

# Logistic model

Let's fit now a logistic regression where we model the outcome, i.e. the positivity
of the covid test, with the following covariates:

- gender

- days since pandemic (our main interest)

- age

- specimen collection via drive-thru site or not 

```{r}
mod <- glm(outcome ~ gender + pan_day + age + drive_thru_ind,
           data = covid_testing, family = "binomial")

tbl_regression(mod, exponentiate = TRUE)
```

We can observed that there's a slight increase in the event odds for age, but
the coefficients of days was not significant. The p-value is slightly larger than the 
usual 0.05, so we cannot completely exclude the existence of a relationship 
between the two variables.

Let's change the time scale: instead of day, we can evaluate weekly associations.

```{r}
covid_testing$pan_week <- floor(covid_testing$pan_day / 7)

mod.week <- glm(outcome ~ gender + pan_week + age + drive_thru_ind,
           data = covid_testing, family = "binomial")

tbl_regression(mod.week, exponentiate = TRUE)
```

Results here do not differ from the previous case. 

# Conclusion

The analyses here do not show a strong relationship between time and covid cases. 
However, these results are not reliable, nor robust, considering that
only 6% of the total sample was tested positive. In this sense, we might not have
enough data to uncover the association between time and covid cases. 

